---
title: "Homework #6"
author: "Sophia Miller"
date: "11/21/2019"
output: github_document
---

```{r setup}
library(tidyverse)
library(modelr)
library(mgcv)
library(patchwork)
set.seed(1)
```

# Problem 1

## Importing and tidying the birthweight dataset:

```{r birthweight_data}
#check missing values
birthweight_data = read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace),
  ) %>% 
  mutate(
    babysex = fct_infreq(babysex),
    frace = fct_infreq(frace),
    malform = fct_infreq(malform),
    mrace = fct_infreq(mrace))

#check for missing data
sum(is.na(birthweight_data))
```

## Model building: 

First, I'll fit univariate models with variables that I believe are associated wih birthweight based on prior knowledge:

```{r univariate}
fit_fincome = lm(bwt ~ fincome, data = birthweight_data) 

fit_gaweeks = lm(bwt ~ gaweeks, data = birthweight_data)

fit_malform = lm(bwt ~ malform, data = birthweight_data)

fit_menarche = lm(bwt ~ menarche, data = birthweight_data)

fit_momage = lm(bwt ~ momage, data = birthweight_data)

fit_mrace = lm(bwt ~ mrace, data = birthweight_data)

fit_ppbmi = lm(bwt ~ ppbmi, data = birthweight_data)

fit_smoken = lm(bwt ~ smoken, data = birthweight_data) 

fit_fincome %>%
  broom::tidy() 

fit_gaweeks %>%
  broom::tidy()

fit_malform %>%
  broom::tidy()

fit_menarche %>%
  broom::tidy() 

fit_momage %>%
  broom::tidy() 

fit_mrace %>%
  broom::tidy() 

fit_ppbmi %>%
  broom::tidy()

fit_smoken %>%
  broom::tidy()
```

Next, I will take variables with a p-value < 0.25 from the bivariate analyses and fit a multivariate model:

```{r multivariable_model}
fit_multi = lm(bwt ~ fincome + gaweeks + momage + mrace + ppbmi + smoken, data = birthweight_data) 

fit_multi %>% 
  broom::tidy()
```

Finally, I will take variables with p-value < 0.05 from the multivariate model to make final model: 

```{r final_model}
#gaweeks, mrace, ppbmi, smoken
fit_final = lm(bwt ~ gaweeks + mrace + ppbmi + smoken, data = birthweight_data)

fit_final %>% 
  broom::tidy()
```

Plot of residuals against fitted values:

```{r residuals_plot}
birthweight_res = modelr::add_residuals(birthweight_data, fit_final) 

birthweight_res_pred = modelr::add_predictions(birthweight_res, fit_final)

ggplot(birthweight_res_pred, aes(x = pred, y = resid)) + geom_point() +
  labs(
    title = "Residuals vs. Fitted Values for Birthweight Model",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

## Model Comparisons

Make comparison models:

* *Model 1 predictors: length at birth and gestational age* 

* *Model 2 predictors: head circumference, length, sex and all interactions* 

```{r comparison_models}
fit_compare = lm(bwt ~ blength + gaweeks, data = birthweight_data)

fit_compare1 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_data)
```

Cross validation:

```{r cv}
cv_df = 
  crossv_mc(birthweight_data, 100) %>% 
  mutate(
    fit_final = map(train, ~lm(bwt ~ gaweeks + mrace + ppbmi + smoken, data = .x)),
    fit_compare = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit_compare1 = map(train, ~lm(bwt ~ bhead + blength + babysex, data = .x))
    ) %>% 
  mutate(
    rmse_final = map2_dbl(fit_final, test, ~rmse(model = .x, data = .y)),
    rmse_compare = map2_dbl(fit_compare, test, ~rmse(model = .x, data = .y)),
    rmse_compare1 = map2_dbl(fit_compare1, test, ~rmse(model = .x, data = .y))
    )
```

Plot prediction error distribution:

```{r prediction_plot}
cv_df %>% 
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
      names_to = "model",
      values_to = "rmse",
      names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
    labs(
    title = "Comparison of Prediction Error Distribution",
    x = "Model",
    y = "RMSE"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
  
```

Based on the plot above, we can see that the third model, which includes head circumference, length, sex and all interactions as predictors is best based on the low RMSE values and shape of the violin plot. The low RMSE values indicate that this model is the best at predicting birthweight.

# Problem 2

Load weather data:

```{r weather_data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Bootstrapping:

```{r bootstrap}
weather_bootstrap = 
  weather_df %>%
  modelr::bootstrap(n = 5000) %>% 
  mutate(models = map(strap, ~lm(tmax ~ tmin, data = .x)))
```

Plotting r2 and beta0*beta1 values:

```{r}
r2_plot = weather_bootstrap %>%
mutate(
  results = map(models, broom::glance)) %>% 
  select(results) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + 
    geom_density() +
    labs(
      title = "Distribution of R^2",
      x = "R-squared",
      y = "Density") +
    theme(plot.title = element_text(hjust = 0.5))

beta_plot = weather_bootstrap %>%
  mutate(
    results = map(models, broom::tidy)) %>% 
  select(results) %>% 
  unnest(results) %>% 
  select(term, estimate) %>% 
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>% 
  unnest() %>% 
  janitor::clean_names() %>% 
  mutate(intercept_beta1 = log(intercept*tmin)) %>% 
    ggplot(aes(x = intercept_beta1)) + 
      geom_density() +
      labs(
        title = "Distribution of log(beta0*beta1)",
        x = "log(beta0*beta1)",
        y = "Density") +
      theme(plot.title = element_text(hjust = 0.5))

r2_plot + beta_plot + plot_layout(widths = c(5,5))
```

The plots show that r-squared and beta0 * beta1 appear to be approximately normally distributed. The mean of r-squared values is about 0.91, while the mean of beta0 * beta1 is about 8.25.


95% confidence interval for r2 and beta0*beta1:

```{r warning = FALSE}
weather_bootstrap %>%
mutate(
  results = map(models, broom::glance)) %>% 
  select(results) %>% 
  unnest(results) %>%
  pull(r.squared) %>% 
  as.vector() %>% 
  quantile(probs = c(0.025, 0.975), na.rm = TRUE)

weather_bootstrap %>%
  mutate(
    results = map(models, broom::tidy)) %>% 
  select(results) %>% 
  unnest(results) %>% 
  select(term, estimate) %>% 
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>% 
  unnest() %>% 
  janitor::clean_names() %>% 
  mutate(intercept_beta1 = intercept + tmin) %>% 
  pull(intercept_beta1) %>% 
  as.vector() %>% 
  quantile(probs = c(0.025, 0.975), na.rm = TRUE)
```

The 95% confidence interval for r-squared values is (0.89, 0.93) and the 95% confidence interval for beta0 * beta1 is (7.80, 8.74).

